---
title: "March Data Hack"
author: "Denys Osmak"
date: "2024-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(sas7bdat)
library(tibble)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
```

# TITLE

## Introduction 

We are here focusing on finding what spacific areas in education are lacking the most. We want to focus what types of schools (Elementry, Middle, High, Elem/Seccondary schools) are relativly behind the avarage preformance. Additionally, what counties are the worst prefformers and the best.

## Methods

### Comparesions of School Types

1.  Find the avarage for all texas schools and the 5 sum summery
2.  Create box plot for all school types (exluding Districts) that will be as a baseline.
3.  Create box plots for all school types seperatly and the avarage plot
4.  Find the worst preforming (bottom quartile and bottom outliers) and do a table on the Counties they are from

### Compersions of Counties 

1.  Find the avarage for all texas schools and the 5 sum summery
2.  Create box plot for all school types (exluding Districts) that will be as a baseline.
3.  Find the worst preforming (bottom quartile and bottom outliers) and do a table on the Counties they are from

## Analysis 

```{r}
#importing data 

schoolData = read.csv("Copy of School_Year_2021-2022_Statewide_Accountability_Ratings_as_of_February_6__2023_20240305.csv")

#Find the avarage for all texas schools and the 5 sum summery 

schoolSum = schoolData %>% 
  filter(School.Type != "District")

notRatedSchools = schoolSum %>% 
  filter(Overall.Score == ".")

ratedSchools = schoolSum %>% 
  filter(Overall.Score != ".")

#Create box plot for all school types (exluding Districts) that will be as a baseline. 

ggplot(ratedSchools, aes(x = School.Type, y = as.integer(Overall.Score))) + geom_boxplot(col="navy")

ratedSchoolsTable = group_by(ratedSchools$School.Type) %>% 
  fivenum(as.integer(ratedSchools$Overall.Score))

ggplot(notRatedSchools, aes(x = School.Type, y = as.integer(Overall.Score))) + geom_boxplot(col="navy")



#Create box plots for all school types seperatly and the avarage plot 

#Find the worst preforming (bottom quartile and bottom outliers) and do a table on the Counties they are from 


```

## Results 

## Conclusion  
